\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-04 paper.qmd
Added images for the anomalies

mnist_proj_pacmap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("b")

@@

mnist_proj_pacmap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.1) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("c")

@@

mnist_proj_pacmap_model3 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("d")

@@

<!--add langevitour screenshots and youtube animation link-->
```{r}
#| label: fig-mnist-tri-proj
#| fig-pos: H
#| fig-height: 10
#| fig-width: 10
#| fig-cap: "Model in \\gD{}, on the PaCMAP layout (a), and three views (b, c, d) of the fit in projections from $10\\text{-}D$, for the digit 1 of MNIST data ($a_1 = 0.062$, $b = 285 \\  (19, \ 15)$, $m = 109$, and benchmark value to remove large edges is $0.108$). (b) The non-linear structure observed in the \\gD{} layout of PaCMAP (@fig-pacmap-author) is also visible when visualizing the model overlaid on the data space. This indicates that PaCMAP accurately captures the non-linear structure of the $10\\text{-}D$ data. (c) the model shows a twisted pattern within the non-linear structure in $10\\text{-}D$ space, which is an additional pattern not visible in the \\gD{} representation (@fig-pacmap-author). (d) some long edges exist in the $10\\text{-}D$ space that are not recognized as long edges in the \\gD{} representation. The full video is available at <https://youtu.be/zcg_GXBmqjA>."

free(trimesh_removed_mnist) + mnist_proj_pacmap_model1 +
  mnist_proj_pacmap_model2 + mnist_proj_pacmap_model3 +
  plot_layout(ncol = 2)
```

@@

img_error_within <- c(6169, 5987, 773, 6671, 4785, 4699, 5173, 5475)

@@

img_error_outside <- c(1489, 5095, 855, 3143, 453, 3619, 1015, 409) #, 5101

@@

```{r}
#| label: fig-model-error-mnist
#| fig-cap: "$10\\text{-}D$ model error in PaCMAP layout of MNIST digit 1 dataset in $2\\text{-}D$. Most low $10\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most large $10\\text{-}D$ model errors are concentrated along the upper edge (a). There are certain data points that exhibit large error rates due to their deviation from the usual $10\\text{-}D$ data structure, which makes them anomalies. These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it. The images associated with high model error points within the non-linear structure display different patterns of the digit 1 (b). However, when comparing these images to the ones found outside of the non-linear structure, it becomes evident that the latter display different patterns of the digit 1 (c). "
#| fig-pos: H
#| fig-width: 4
#| fig-height: 8
#| out-height: 80%

>>>>

mnist_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("b")

@@

mnist_proj_tsne_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.1) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("c")

@@

mnist_proj_tsne_model3 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("d")

@@

<!--add langevitour screenshots and youtube animation link-->
```{r}
#| label: fig-mnist-tri-proj
#| fig-pos: H
#| fig-height: 10
#| fig-width: 10
#| fig-cap: "Model in \\gD{}, on the tSNE layout (a), and three views (b, c, d) of the fit in projections from $10\\text{-}D$, for the digit 1 of MNIST data ($a_1 = 0.062$, $b = 285 \\  (19, \ 15)$, $m = 109$, and benchmark value to remove large edges is $0.108$). (b) The non-linear structure observed in the \\gD{} layout of PaCMAP (@fig-pacmap-author) is also visible when visualizing the model overlaid on the data space. This indicates that PaCMAP accurately captures the non-linear structure of the $10\\text{-}D$ data. (c) the model shows a twisted pattern within the non-linear structure in $10\\text{-}D$ space, which is an additional pattern not visible in the \\gD{} representation (@fig-pacmap-author). (d) some long edges exist in the $10\\text{-}D$ space that are not recognized as long edges in the \\gD{} representation. The full video is available at <https://youtu.be/zcg_GXBmqjA>."

free(trimesh_removed_mnist) + mnist_proj_tsne_model1 +
  mnist_proj_tsne_model2 + mnist_proj_tsne_model3 +
  plot_layout(ncol = 2)
```

@@

img_error_within <- c(5173, 6671, 2964, 2464, 6319, 7388, 4303, 7650)

@@

img_error_outside <- c(3554, 501, 4422, 6860, 3869, 1529, 4916, 6144)

@@

```{r}
#| label: fig-model-error-mnist
#| fig-cap: "$10\\text{-}D$ model error in tSNE layout of MNIST digit 1 dataset in $2\\text{-}D$. Most low $10\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most large $10\\text{-}D$ model errors are concentrated along the upper edge (a). There are certain data points that exhibit large error rates due to their deviation from the usual $10\\text{-}D$ data structure, which makes them anomalies. These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it. The images associated with high model error points within the non-linear structure display different patterns of the digit 1 (b). However, when comparing these images to the ones found outside of the non-linear structure, it becomes evident that the latter display different patterns of the digit 1 (c). "
#| fig-pos: H
#| fig-width: 4
#| fig-height: 8
#| out-height: 80%

### 2024-09-04 paper.tex
Added five layouts for MNIST
Added error plot for MNSIT data
Try to tune with different hyper parameter choices, but nothing much changed with default settings. Therefore, the best with tSNE
Added images for the anomalies

@@

\caption{\label{fig-mnist-tri-proj}Model in \gD{}, on the PaCMAP layout
(a), and three views (b, c, d) of the fit in projections from
\(10\text{-}D\), for the digit 1 of MNIST data (\(a_1 = 0.062\),
\(b = 285 \  (19,  15)\), \(m = 109\), and benchmark value to remove
large edges is \(0.108\)). (b) The non-linear structure observed in the
\gD{} layout of PaCMAP (Figure~\ref{fig-pacmap-author}) is also visible
when visualizing the model overlaid on the data space. This indicates
that PaCMAP accurately captures the non-linear structure of the
\(10\text{-}D\) data. (c) the model shows a twisted pattern within the
non-linear structure in \(10\text{-}D\) space, which is an additional
pattern not visible in the \gD{} representation
(Figure~\ref{fig-pacmap-author}). (d) some long edges exist in the
\(10\text{-}D\) space that are not recognized as long edges in the \gD{}
representation. The full video is available at
\url{https://youtu.be/zcg_GXBmqjA}.}

@@

\caption{\label{fig-model-error-mnist}\(10\text{-}D\) model error in
PaCMAP layout of MNIST digit 1 dataset in \(2\text{-}D\). Most low
\(10\text{-}D\) model errors are distributed along the lower edge of the
\(2\text{-}D\) structure, while most large \(10\text{-}D\) model errors
are concentrated along the upper edge (a). There are certain data points
that exhibit large error rates due to their deviation from the usual
\(10\text{-}D\) data structure, which makes them anomalies. These
anomalies can be classified into two types: those that are anomalies
within the non-linear structure and those that lie outside of it. The
images associated with high model error points within the non-linear
structure display different patterns of the digit 1 (b). However, when
comparing these images to the ones found outside of the non-linear
structure, it becomes evident that the latter display different patterns
of the digit 1 (c).}

>>>>

\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-mnist-mse-1.pdf}

\caption{\label{fig-mnist-mse}Different methods with default
hyper-parameter settings}

@@

\caption{\label{fig-mnist-tri-proj}Model in \gD{}, on the tSNE layout
(a), and three views (b, c, d) of the fit in projections from
\(10\text{-}D\), for the digit 1 of MNIST data (\(a_1 = 0.062\),
\(b = 285 \  (19,  15)\), \(m = 109\), and benchmark value to remove
large edges is \(0.108\)). (b) The non-linear structure observed in the
\gD{} layout of PaCMAP (Figure~\ref{fig-pacmap-author}) is also visible
when visualizing the model overlaid on the data space. This indicates
that PaCMAP accurately captures the non-linear structure of the
\(10\text{-}D\) data. (c) the model shows a twisted pattern within the
non-linear structure in \(10\text{-}D\) space, which is an additional
pattern not visible in the \gD{} representation
(Figure~\ref{fig-pacmap-author}). (d) some long edges exist in the
\(10\text{-}D\) space that are not recognized as long edges in the \gD{}
representation. The full video is available at
\url{https://youtu.be/zcg_GXBmqjA}.}

@@

\caption{\label{fig-model-error-mnist}\(10\text{-}D\) model error in
tSNE layout of MNIST digit 1 dataset in \(2\text{-}D\). Most low
\(10\text{-}D\) model errors are distributed along the lower edge of the
\(2\text{-}D\) structure, while most large \(10\text{-}D\) model errors
are concentrated along the upper edge (a). There are certain data points
that exhibit large error rates due to their deviation from the usual
\(10\text{-}D\) data structure, which makes them anomalies. These
anomalies can be classified into two types: those that are anomalies
within the non-linear structure and those that lie outside of it. The
images associated with high model error points within the non-linear
structure display different patterns of the digit 1 (b). However, when
comparing these images to the ones found outside of the non-linear
structure, it becomes evident that the latter display different patterns
of the digit 1 (c).}
